\todo{New title} Concretizing the Process Hitting into Biological Regulatory Networks

### 2013-04-29 parts/infer-param.tex
continue refactoring section 4.3

\newcommand{\paramlabel}{\atom{param\_label}}
\newcommand{\paramres}{\atom{param\_resource}}

@@

A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
For instance, all information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
In other words, $\paramlabel(a, p)$ states that a parameter $K^p_{a,\omega}$ exists.
Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 7 parameters, and to state the existence of each of them, the following facts are used:
\begin{align*}
  &\paramlabel(a,1).
  &\paramlabel(b,1).\\
  &\paramlabel(a,2).
  &\paramlabel(b,2).\\
  &\paramlabel(a,3).\\
  &\paramlabel(a,4).
  &\paramlabel(c,1).
\end{align*}

>>>>

\newcommand{\paramlabel}{\atom{param\_label}}
\newcommand{\paramres}{\atom{param\_resource}}
\newcommand{\component}{\atom{component}}
\newcommand{\componentlevels}{\atom{component\_levels}}

@@

\rewrite{Facts : component}

\rewrite{Rules + variables : component\_levels}

\rewrite{Cardinalities : énumération de param + param\_label et infered\_param}

A rule with no $body$ part is called a fact, and its $head$ atom has to belong to all answer sets.
For instance, the information describing the studied model (PH and inferred IG \& parameters) are expressed in ASP using facts.
In particular, the predicate $\component$ allows to define all components belonging to the inferred IG.
Thus, an atom $\component(a, m)$ states that $a$ is a component of the IG and $l_a = m$.
Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 3 components, and to state the existence of each of them, the following facts are used:
\begin{align*}
  &\component(a, 2).
  &\component(c, 1). \\
  &\component(b, 1).
\end{align*}

To define the sets of possible expression levels of each component (\ie the set $\segm{0}{l_a}$ for each $a \in \Gamma$),
one can use atoms of the form $\componentlevels(a, k)$ to state that $k \in \segm{0}{l_a}$.
Variables are then useful in order to enumerate each possible $k$ for all component $a$.
A variable starts with a capital letter, and during the solving,
any rule containing variables will be duplicated in order to replace each variable by all the possible values it could represent.
\rewrite{Reformulate}
For example, the following rule, which contains three variables ($\var{A}$, $\var{K}$ and $\var{M}$), enumerates all possible expression levels of all components in the system:
\begin{align*}
  \componentlevels(A, K) \la \component(\var{A}, \var{M}), 0 \leq K \leq M.
\end{align*}
The “$\leq$” notation stand for a shortcut in ASP which has the same meaning as the mathematical operator.
Regarding \pref{ex:infer-param-runningPH-1}, the previous rule will make the following set of atoms belong to all answer sets:
\begin{align*}
  &\{&\componentlevels(b, 0)
  &&\componentlevels(b, 1) \\
  &&\componentlevels(c, 0)
  &&\componentlevels(c, 1) \\
  &&\componentlevels(a, 0)
  &&\componentlevels(a, 1) \\
  &&\componentlevels(a, 2) &&&&\}
\end{align*}

Cardinality rules are convenient to enumerate all possible parametrizations by creating multiple answer sets.
The head of a cardinality rule specifies a set of atoms $H$ and two integers $min$ and $max$, and is denoted:
\begin{align*}
  min\ \{\ H\ \}\ max \la \atom{A}_1, \atom{A}_2, \dots, \atom{A}_n, \nota \atom{B}_1, \nota \atom{B}_2, \dots, \nota \atom{B}_m.
\end{align*}
Given such a rule, every answer $S$ sets have to verify:
\begin{align*}
  min \leq |S \cap H| \leq max
\end{align*}
In other words, all answer sets contain a subset of $H$ whose cardinality goes from $min$ to $max$.
The set of atoms $H = \{ \atom{H}_1, \atom{H}_2, \dots, \atom{H}_p \}$ is often defined as: $H = \atom{P} : \atom{Q}$,
which is a shorthand for “all atoms of the form $\atom{P}$ for which $\atom{Q}$ is true”.
\rewrite{Reformulate}

Given a set of atoms, it creates as many answer sets as there are possible 
constrains gives any number of possible answers for some atoms between a lower and upper bounds.
For example,
\begin{align*}
  & 1\ \{ \atom{param}(\var{X}, \var{P}, \var{I}) : \atom{component\_levels}(\var{X}, \var{I}) \} \la \\
  & \qquad\qquad\qquad \paramlabel(\var{X}, \var{P}), \nota \atom{infered\_param}(\var{X}, \var{P}).
\end{align*}

---------------

In particular, the existence of a parameter (whether its value has been inferred or not) is stated by the atom $\paramlabel$.
For functional purposes, we assign a unique label to every possible set of resources $\omega$ of a given component,
and in the following we note $K^p_{a,\omega}$ the parameter of component $a$ whose set of resources $\omega$ is assigned to the label $p$.
In other words, $\paramlabel(a, p)$ states that a parameter $K^p_{a,\omega}$ exists.
Consider \pref{ex:infer-param-runningPH-1}: the inferred IG contains 7 parameters, and to state the existence of each of them, the following facts are used:
\begin{align*}
  &\paramlabel(a,1).
  &\paramlabel(b,1).\\
  &\paramlabel(a,2).
  &\paramlabel(b,2).\\
  &\paramlabel(a,3).\\
  &\paramlabel(a,4).
  &\paramlabel(c,1).
\end{align*}
